---
title: "presentation ready plots"
date: '2021-09-29'
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: solarized-light
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
      beforeInit: "https://platform.twitter.com/widgets.js"
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
xaringanExtra::use_panelset()
library(countdown)
```

```{r xaringan-themer, include = FALSE, warning = FALSE, fig.showtext = FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#866fa3",
  secondary_color = "#F1DE67",
  inverse_header_color = "#464a53",
  black_color = "#464a53",
  code_highlight_color = "#f1de67",
  header_font_google = google_font("Atkinson Hyperlegible"),
  text_font_google   = google_font("Atkinson Hyperlegible", "300", "300i"),
  code_font_google   = google_font("Source Code Pro"),
  code_font_size = "20px",
  title_slide_background_color = "#FFFFFF",
  title_slide_background_image = "https://github.com/vizdata-f21/slides/raw/main/vizdata-bg.jpeg",
  title_slide_background_size = "contain",
  base_font_size = "24px",
  header_h1_font_size = "1.9rem",
  header_h2_font_size = "1.75rem",
  header_h3_font_size = "1.5rem",
  extra_css = list(
    "h1" = list("margin-block-start" = "0.4rem", 
                 "margin-block-end" = "0.4rem"),
    "h2" = list("margin-block-start" = "0.4rem", 
                 "margin-block-end" = "0.4rem"),
    "h3" = list("margin-block-start" = "0.4rem", 
                 "margin-block-end" = "0.4rem"),
    ".tiny" = list("font-size" = "70%"),
    ".small" = list("font-size" = "80%"),
    ".midi" = list("font-size" = "90%"),
    ".tiny .remark-code" = list("font-size" = "70%"),
    ".small .remark-code" = list("font-size" = "80%"),
    ".midi .remark-code" = list("font-size" = "90%"),
    ".large" = list("font-size" = "200%"),
    ".huge" = list("font-size" = "400%",
                     "font-family" = "'Montserrat', sans-serif",
                     "font-weight" = "bold"),
    ".hand" = list("font-family" = "'Gochi Hand', cursive",
                   "font-size" = "125%"),
    ".task" = list("padding-right"    = "10px",
                   "padding-left"     = "10px",
                   "padding-top"      = "3px",
                   "padding-bottom"   = "3px",
                   "margin-bottom"    = "6px",
                   "margin-top"       = "6px",
                   "border-left"      = "solid 5px #F1DE67",
                   "background-color" = "#F1DE6750"),
    ".pull-left" = list("width" = "49%",
                        "float" = "left"),
    ".pull-right" = list("width" = "49%",
                         "float" = "right"),
    ".pull-left-wide" = list("width" = "70%",
                             "float" = "left"),
    ".pull-right-narrow" = list("width" = "27%",
                                "float" = "right"),
    ".pull-left-narrow" = list("width" = "27%",
                               "float" = "left"),
    ".pull-right-wide" = list("width" = "70%",
                              "float" = "right"),
    ".blue" = list(color = "#2A9BB7"),
    ".purple" = list(color = "#a493ba"),
    ".yellow" = list(color = "#f1de67"),
    ".gray" = list(color = "#464a53")
    )
  )
```

class: middle, inverse

# Welcome

---

## Announcements

- Project presentations on Friday
  - Bring a laptop to fill out feedback forms
  - One more team evaluation to be filled out promptly after class
  - Presentation order, with latest start-by times:
     - 12:05 - Ctrl+Alt+Elite
     - 12:15 - Team 5 Look Alive
     - 12:25 - Mr. Palmer's Penguins
     - 12:35 - R-mageddon
     - 12:45 - 7 Up
     - 12:55 - The Outliers
     - 13:05 - Dorian's Gs
     - 13:15 - The Scatterplots

- After fall break:
  - Reading quiz 3 due Wed, Oct 6
  - New teams!
  - Take note of Project 2 presentation date: must be in class in person to present or present live via Zoom in the case of isolation

---

## Setup

.midi[
```{r message = FALSE, warning = FALSE}
# load packages
library(tidyverse)
library(colorspace)
library(ggtext)
library(scales)

# set default theme for ggplot2
ggplot2::theme_set(ggplot2::theme_minimal(base_size = 16))

# set default figure parameters for knitr
knitr::opts_chunk$set(
  fig.width = 8, fig.asp = 0.618, fig.retina = 3,
  dpi = 300, out.width = "60%"
)

# dplyr print min and max
options(dplyr.print_max = 6, dplyr.print_min = 6)
```
]

---

## Agenda for today

- Midterm feedback summary

- Wrap up: Annotation

- Presentation ready plots

---

class: inverse, middle

# Midterm feedback summary

---

## An opportunity to talk more dataviz?

```{r}
midterm_eval <- tibble::tribble(
    ~class_pace,                ~class_time,
  "About right",             "More lecture",
  "About right",         "More live coding",
  "About right",         "More live coding",
  "About right",         "More live coding",
  "About right",         "More live coding",
  "About right",         "More live coding",
  "About right",    "More working in teams",
  "About right", "No change, keep it as is",
  "About right", "No change, keep it as is",
  "About right", "No change, keep it as is",
     "Too fast",         "More live coding",
     "Too fast", "No change, keep it as is",
     "Too slow",         "More live coding"
  )
```

---

## Class pace 1

.pull-left[
```{r class-pace-1, fig.show = "hide"}
ggplot(midterm_eval, aes(x = class_pace)) +
  geom_bar()
```
]
.pull-right[
```{r ref.label = "class-pace-1", echo = FALSE, out.width = "100%"}
```
]

---

## Class pace 2

.pull-left[
```{r class-pace-2, fig.show = "hide"}
ggplot(midterm_eval, aes(x = class_pace)) +
  geom_bar() +
  scale_y_continuous(breaks = seq(0, 10, 2)) #<<
```
]
.pull-right[
```{r ref.label = "class-pace-2", echo = FALSE, out.width = "100%"}
```
]

---

## Class pace 3

.pull-left[
```{r class-pace-3, fig.show = "hide"}
ggplot(midterm_eval, 
       aes(y = factor(1), #<<
           fill = class_pace) #<<
       ) +
  geom_bar(position = "fill") #<<
```
]
.pull-right[
```{r ref.label = "class-pace-3", echo = FALSE, out.width = "100%"}
```
]

---

## Class pace 4

.pull-left[
```{r class-pace-4, fig.show = "hide"}
midterm_eval %>%
  mutate( #<<
    class_pace = fct_relevel(class_pace, #<<
                             "Too fast", #<<
                             "About right", #<<
                             "Too slow") #<<
    ) %>% #<<
  ggplot(aes(y = factor(1), fill = class_pace)) +
  geom_bar(position = "fill") 
```
]
.pull-right[
```{r ref.label = "class-pace-4", echo = FALSE, out.width = "100%"}
```
]

---

## Class pace 5

.pull-left[
```{r class-pace-5, fig.show = "hide"}
midterm_eval %>%
  mutate(class_pace = fct_relevel(class_pace, "Too fast", "About right", "Too slow")) %>% 
  ggplot(
    aes(y = "Class pace", #<<
        fill = class_pace)
    ) +
  geom_bar(position = "fill") 
```
]
.pull-right[
```{r ref.label = "class-pace-5", echo = FALSE, out.width = "100%"}
```
]

---

## Class pace 6

.pull-left[
```{r class-pace-6, fig.show = "hide"}
midterm_eval %>%
  mutate(class_pace = fct_relevel(class_pace, "Too fast", "About right", "Too slow")) %>% 
  ggplot(aes(y = "Class pace", fill = class_pace)) +
  geom_bar(position = "fill") +
  labs( #<<
    x = "Percentage", #<<
    y = NULL, #<<
    fill = NULL #<<
    ) #<<
```
]
.pull-right[
```{r ref.label = "class-pace-6", echo = FALSE, out.width = "100%"}
```
]

---

## Class pace 7

.pull-left[
```{r class-pace-7, fig.show = "hide"}
midterm_eval %>%
  mutate(class_pace = fct_relevel(class_pace, "Too fast", "About right", "Too slow")) %>% 
  ggplot(aes(y = "Class pace", fill = class_pace)) +
  geom_bar(position = "fill") +
  labs(x = "Percentage", y = NULL, fill = NULL) +
  theme(legend.position = "bottom") #<<
```
]
.pull-right[
```{r ref.label = "class-pace-7", echo = FALSE, out.width = "100%"}
```
]

---

## Class pace 8

.pull-left[
```{r class-pace-8, fig.show = "hide"}
midterm_eval %>%
  mutate(class_pace = fct_relevel(class_pace, "Too fast", "About right", "Too slow")) %>% 
  ggplot(aes(y = "Class pace", fill = class_pace)) +
  geom_bar(position = "fill") +
  labs(x = "Percentage", y = NULL, fill = NULL) +
  theme(legend.position = "bottom") +
  scale_fill_discrete( #<<
    guide = guide_legend(reverse = TRUE) #<<
    ) #<<
```
]
.pull-right[
```{r ref.label = "class-pace-8", echo = FALSE, out.width = "100%"}
```
]

---

## Class pace 9

.pull-left[
```{r class-pace-9, fig.show = "hide"}
midterm_eval %>%
  mutate(class_pace = fct_relevel(class_pace, "Too fast", "About right", "Too slow")) %>% 
  ggplot(aes(y = "Class pace", fill = class_pace)) +
  geom_bar(position = "fill") +
  labs(x = "Percentage", y = NULL, fill = NULL) +
  theme(legend.position = "bottom") +
  scale_fill_discrete(guide = guide_legend(reverse = TRUE)) +
  scale_x_continuous( #<<
    labels = label_percent() #<<
    ) #<<
```
]
.pull-right[
```{r ref.label = "class-pace-9", echo = FALSE, out.width = "100%"}
```
]

---

## Class pace 10

.pull-left[
```{r class-pace-10, fig.show = "hide"}
midterm_eval %>%
  mutate(class_pace = fct_relevel(class_pace, "Too fast", "About right", "Too slow")) %>% 
  ggplot(aes(y = "Class pace", fill = class_pace)) +
  geom_bar(position = "fill") +
  labs(x = "Percentage", y = NULL, fill = NULL) +
  theme(legend.position = "bottom") +
  scale_fill_discrete_diverging( #<<
    guide = guide_legend(reverse = TRUE)
    ) + 
  scale_x_continuous(labels = label_percent())
```
]
.pull-right[
```{r ref.label = "class-pace-10", echo = FALSE, out.width = "100%"}
```
]

---

## Class pace 11

.pull-left[
```{r class-pace-11, fig.show = "hide"}
midterm_eval %>%
  mutate(class_pace = fct_relevel(class_pace, "Too fast", "About right", "Too slow")) %>% 
  ggplot(aes(y = "Class pace", fill = class_pace)) +
  geom_bar(position = "fill") +
  labs(x = "Percentage", y = NULL, fill = NULL) +
  theme(legend.position = "bottom") +
  scale_fill_discrete_diverging(guide = guide_legend(reverse = TRUE)) +
  scale_x_continuous(
    labels = label_percent(), 
    position = "top" #<<
    )
```
]
.pull-right[
```{r ref.label = "class-pace-11", echo = FALSE, out.width = "100%"}
```
]

---

## Class pace 12

No change in code, only adjust `fig.asp` to 0.4.

.pull-left[
```{r class-pace-12, fig.show = "hide"}
midterm_eval %>%
  mutate(class_pace = fct_relevel(class_pace, "Too fast", "About right", "Too slow")) %>% 
  ggplot(aes(y = "Class pace", fill = class_pace)) +
  geom_bar(position = "fill") +
  labs(x = "Percentage", y = NULL, fill = NULL) +
  theme(legend.position = "bottom") +
  scale_fill_discrete_diverging(guide = guide_legend(reverse = TRUE)) +
  scale_x_continuous(labels = label_percent(), position = "top" )
```
]
.pull-right[
```{r ref.label = "class-pace-12", echo = FALSE, out.width = "100%", fig.asp = 0.4}
```
]

---

## Class time

.panelset.sideways[
```{r panelset = c(source = "Code", output = "Plot"), out.width = "100%", fig.asp = 0.4}
midterm_eval %>%
  mutate(
    class_time = str_remove(class_time, ", keep it as is"),
    class_time = fct_relevel(class_time, "More lecture", "No change", "More live coding", "More working in teams")
    ) %>%
  ggplot(aes(y = "Class time", fill = class_time)) +
  geom_bar(position = "fill") +
  labs(
    x = "Percentage",
    y = NULL,
    fill = NULL
  ) +
  theme(
    legend.position = "bottom",
    legend.direction = "vertical"
    ) +
  scale_fill_manual(
    values = c("#BCC1D4", "#E2E2E2", "#D6BCC0", "#8E063B"), 
    guide = guide_legend(nrow = 2, reverse = TRUE)
    ) +
  scale_x_continuous(labels = label_percent(), position = "top")
```
]

---

## Other feedback

- Need more practice:
  - Scaling
  - Data cleaning/transformation
  - Planning a unique visualization

--
- Time spend: 
  - ~6 hrs/week on average out of class
  - 3.75 hrs/week in class
  - Total: ~9.75 hrs/week

--
- Misc:
  - Grading: 
    - HW: Randomly assign to graders (TAs) for each assignment
    - Project: Averaged across all graders (TAs + prof)
  - Release assignments earlier
  - Tips for creating a web (GitHub) portfolio

---

class: inverse, middle

# Wrap up: Annotation

---

## Explanations -> annotations

**Rule of thumb:** If you feel the need to draw your audience's attention to a particular feature of the plot, experiment with doing so using annotation instead of in words.

---

## Rewind: Sizes of planes

```{r include = FALSE}
rdu_flights_2020 <- read_csv(here::here("10-presentation-plots", "data/rdu-flights-2020.csv"))
rdu_planes_2020 <- read_csv(here::here("10-presentation-plots", "data/rdu-planes-2020.csv"))
```

```{r eval = FALSE}
rdu_flights_2020 <- read_csv(here::here("data/rdu-flights-2020.csv"))
rdu_planes_2020 <- read_csv(here::here("data/rdu-planes-2020.csv"))
```

```{r}
rdu_flights_planes <- rdu_flights_2020 %>%
  inner_join(rdu_planes_2020, by = "tailnum") %>%
  mutate(
    size = case_when(
      seats <= 30 ~ "Small",
      seats >= 31 & seats <= 110 ~ "Medium",
      seats >= 111 & seats <= 210 ~ "Large",
      seats > 211 ~ "Jumbo"
    ),
    size = fct_relevel(size, "Small", "Medium", "Large", "Jumbo")
  )
```

---

.task[
How can we improve this visualization, in terms of color, axes, and annotation?
]

.panelset.sideways[
```{r panelset = c(source = "Code", output = "Plot"), out.width = "100%"}
rdu_flights_planes %>%
  count(hour, size) %>%
  ggplot(aes(x = hour, y = n, color = size)) +
  geom_line() +
  labs(
    x = "Hour of day",
    y = "Number of flights",
    color = "Size of plane",
    title = "Number of flights out of RDU in 2020",
    subtitle = "By size of plane",
    caption = "Source: FAA Aircraft Registry and\nBureau of Transportation Statistics"
  )
```
]

---

class: middle

.hand[livecoding...]

---
