---
title: "interactive<br>visualization"
date: '2021-11-11'
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: solarized-light
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
      beforeInit: "https://platform.twitter.com/widgets.js"
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
xaringanExtra::use_panelset()
library(countdown)
```

```{r xaringan-themer, include = FALSE, warning = FALSE, fig.showtext = FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#866fa3", # purple
  secondary_color = "#F1DE67", # yellow
  inverse_header_color = "#464a53", # dark gray
  black_color = "#464a53",
  code_highlight_color = "#f1de67",
  header_font_google = google_font("Atkinson Hyperlegible"),
  text_font_google   = google_font("Atkinson Hyperlegible"),
  code_font_google   = google_font("Source Code Pro"),
  code_font_size = "20px",
  title_slide_background_color = "#FFFFFF",
  title_slide_background_image = "https://github.com/vizdata-f21/slides/raw/main/vizdata-bg.jpeg",
  title_slide_background_size = "contain",
  base_font_size = "24px",
  header_h1_font_size = "1.9rem",
  header_h2_font_size = "1.75rem",
  header_h3_font_size = "1.5rem",
  extra_css = list(
    "h1" = list("margin-block-start" = "0.4rem", 
                 "margin-block-end" = "0.4rem"),
    "h2" = list("margin-block-start" = "0.4rem", 
                 "margin-block-end" = "0.4rem"),
    "h3" = list("margin-block-start" = "0.4rem", 
                 "margin-block-end" = "0.4rem"),
    ".tiny" = list("font-size" = "50%"),
    ".xsmall" = list("font-size" = "72%"),
    ".small" = list("font-size" = "80%"),
    ".midi" = list("font-size" = "90%"),
    ".tiny .remark-code" = list("font-size" = "50%"),
    ".xsmall .remark-code" = list("font-size" = "72%"),
    ".small .remark-code" = list("font-size" = "80%"),
    ".midi .remark-code" = list("font-size" = "90%"),
    ".large" = list("font-size" = "200%"),
    ".huge" = list("font-size" = "400%",
                     "font-family" = "'Montserrat', sans-serif",
                     "font-weight" = "bold"),
    ".hand" = list("font-family" = "'Gochi Hand', cursive",
                   "font-size" = "125%"),
    ".task" = list("padding-right"    = "10px",
                   "padding-left"     = "10px",
                   "padding-top"      = "3px",
                   "padding-bottom"   = "3px",
                   "margin-bottom"    = "6px",
                   "margin-top"       = "6px",
                   "border-left"      = "solid 5px #F1DE67",
                   "background-color" = "#F1DE6750"),
    ".note" = list("padding-right"    = "10px",
               "padding-left"     = "10px",
               "padding-top"      = "3px",
               "padding-bottom"   = "3px",
               "margin-bottom"    = "6px",
               "margin-top"       = "6px",
               "border-left"      = "solid 5px #866fa3",
               "background-color" = "#866fa350"),
    ".pull-left" = list("width" = "49%",
                        "float" = "left"),
    ".pull-right" = list("width" = "49%",
                         "float" = "right"),
    ".pull-left-wide" = list("width" = "70%",
                             "float" = "left"),
    ".pull-right-narrow" = list("width" = "27%",
                                "float" = "right"),
    ".pull-left-narrow" = list("width" = "27%",
                               "float" = "left"),
    ".pull-right-wide" = list("width" = "70%",
                              "float" = "right"),
    ".blue" = list(color = "#2A9BB7"),
    ".purple" = list(color = "#a493ba"),
    ".yellow" = list(color = "#f1de67"),
    ".gray" = list(color = "#464a53")
    )
  )
```

class: middle, inverse

# Welcome

---

## Announcements

- Peer review of Project 2 proposals on Monday in lab

- HW 5 posted

---

## Interactive visualizations with Shiny

We will build an app with

- Data pre-processing
- Dynamic UI elements
- Linked brushing

---

class: center, middle

.midi[
[https://minecr.shinyapps.io/manager-survey](https://minecr.shinyapps.io/manager-survey/)
]

```{r echo = FALSE}
knitr::include_app("https://minecr.shinyapps.io/manager-survey/", height ="550px")
```

---

class: middle

.large[.hand[livecoding]]

.task[
Go to the `exercise-08` project and code along in `manager-survey/app.R`.
]

(See next slide for code developed during live coding session)

---

class: middle

The code for the app can be found [here]().
