---
title: "wrap up"
date: "2021-11-30"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: solarized-light
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
      beforeInit: "https://platform.twitter.com/widgets.js"
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
xaringanExtra::use_panelset()
library(countdown)
```

```{r xaringan-themer, include = FALSE, warning = FALSE, fig.showtext = FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#866fa3", # purple
  secondary_color = "#F1DE67", # yellow
  inverse_header_color = "#464a53", # dark gray
  black_color = "#464a53",
  code_highlight_color = "#f1de67",
  header_font_google = google_font("Atkinson Hyperlegible"),
  text_font_google = google_font("Atkinson Hyperlegible"),
  code_font_google = google_font("Source Code Pro"),
  code_font_size = "20px",
  title_slide_background_color = "#FFFFFF",
  title_slide_background_image = "https://github.com/vizdata-f21/slides/raw/main/vizdata-bg.jpeg",
  title_slide_background_size = "contain",
  base_font_size = "24px",
  header_h1_font_size = "1.9rem",
  header_h2_font_size = "1.75rem",
  header_h3_font_size = "1.5rem",
  extra_css = list(
    "h1" = list(
      "margin-block-start" = "0.4rem",
      "margin-block-end" = "0.4rem"
    ),
    "h2" = list(
      "margin-block-start" = "0.4rem",
      "margin-block-end" = "0.4rem"
    ),
    "h3" = list(
      "margin-block-start" = "0.4rem",
      "margin-block-end" = "0.4rem"
    ),
    ".tiny" = list("font-size" = "50%"),
    ".xsmall" = list("font-size" = "72%"),
    ".small" = list("font-size" = "80%"),
    ".midi" = list("font-size" = "90%"),
    ".tiny .remark-code" = list("font-size" = "50%"),
    ".xsmall .remark-code" = list("font-size" = "72%"),
    ".small .remark-code" = list("font-size" = "80%"),
    ".midi .remark-code" = list("font-size" = "90%"),
    ".large" = list("font-size" = "200%"),
    ".huge" = list(
      "font-size" = "400%",
      "font-family" = "'Montserrat', sans-serif",
      "font-weight" = "bold"
    ),
    ".hand" = list(
      "font-family" = "'Gochi Hand', cursive",
      "font-size" = "125%"
    ),
    ".task" = list(
      "padding-right" = "10px",
      "padding-left" = "10px",
      "padding-top" = "3px",
      "padding-bottom" = "3px",
      "margin-bottom" = "6px",
      "margin-top" = "6px",
      "border-left" = "solid 5px #F1DE67",
      "background-color" = "#F1DE6750"
    ),
    ".note" = list(
      "padding-right" = "10px",
      "padding-left" = "10px",
      "padding-top" = "3px",
      "padding-bottom" = "3px",
      "margin-bottom" = "6px",
      "margin-top" = "6px",
      "border-left" = "solid 5px #866fa3",
      "background-color" = "#866fa350"
    ),
    ".pull-left" = list(
      "width" = "49%",
      "float" = "left"
    ),
    ".pull-right" = list(
      "width" = "49%",
      "float" = "right"
    ),
    ".pull-left-wide" = list(
      "width" = "70%",
      "float" = "left"
    ),
    ".pull-right-narrow" = list(
      "width" = "27%",
      "float" = "right"
    ),
    ".pull-left-narrow" = list(
      "width" = "27%",
      "float" = "left"
    ),
    ".pull-right-wide" = list(
      "width" = "70%",
      "float" = "right"
    ),
    ".blue" = list(color = "#2A9BB7"),
    ".purple" = list(color = "#a493ba"),
    ".yellow" = list(color = "#f1de67"),
    ".gray" = list(color = "#464a53")
  )
)
```

class: middle, inverse

# Welcome

---

## Announcements

- Project presentations on Friday
  - Presentation order to be announced in class on Friday
  - We might run over by a few minutes
  - Bring laptops, submit evaluations during presentations

--
- Remaining office hours:
  - Jennifer: 3-4pm on Wednesday and 5-6pm on Thursday (both virtual)
  - Mine: 3:30-5pm on Thursday (virtual)

--
- Remaining to do items:
  - Project 2
  - Peer evaluations (by 5pm on Friday)

--
- Course and TA evaluations (let's take some time now)

---

class: middle, inverse

# Workshop and peer review

---

class: middle

.task[
Any questions / things I can clarify about the projects or anything else we have covered so far?
]

---

## Peer review - Round 1

- Sit with your team
- Review the project homepage of one of the teams whose repo you have access to
- Open at least 1 issue providing feedback on something you like
- Open at least 1 issue providing a suggestion for improvement

```{r echo = FALSE}
countdown(minutes = 5)
```

---

## Peer review - Round 2

- Sit with your team
- Review the project homepage of one of the teams whose repo you have access to
- Open at least 1 issue providing feedback on something you like
- Open at least 1 issue providing a suggestion for improvement

```{r echo = FALSE}
countdown(minutes = 5)
```

---

## Peer review - Reflection

- Read the issues opened on your repos
- Close the issue with positive feedback with a thank you
- Make a plan for addressing the issue with constructive feedback (or, if minor, handle it now and close with a commit)

```{r echo = FALSE}
countdown(minutes = 5)
```

---

class: middle, inverse

# ...

---

class: middle, inverse

# AMA with Thomas Lin Pedersen

---

class: middle, inverse

# THANK YOU!
